---
title: "Batch sensitivity analysis"
output: html_document
---

Load packages and data.
The data are summary metrics computed by batchSummary.R
Remove runs that resulted in no species.

```{r, message = FALSE}
library(tidyverse)
library(shapper)
library(gridExtra)

data <- read_csv("Y:/TKeggin/genesis/v1.0/output/1d_2000m_17c/5_all/summary_table.csv")
data <- data %>% filter(species.surviving > 0)
```

Explore relationships graphically between start parameters and the simulation's discrepancy with observed patterns of species richness.
Linear regressions are also fitted to each parameter comparison.

```{r, echo = FALSE}
parameters <- c("t_opt","dispRange","speciation","initialAbundance")


for(parameter in parameters){
  plotData <- data.frame(data$richness.discrepancy,data[,parameter])
  
  regression <- lm(plotData[,2] ~ plotData[,1])
  
  print(ggplot(data, aes(x=richness.discrepancy, y=plotData[,2]))+
  labs(y=parameter, x="richness discrepancy") +scale_fill_viridis_c() +
  geom_point(aes(fill=species.surviving),colour = "black", shape =21, size = 2) +
    geom_abline(slope = regression$coefficients[2],
                intercept = regression$coefficients[1]) +
  theme_classic()
  )
}

print(ggplot(data, aes(x=richness.discrepancy, y=species.surviving))+
        geom_point() +
labs(y="species.surviving", x="richness discrepancy") +
scale_colour_viridis_c() +
theme_classic()
)
```


Explore combinations of parameters.
First, need to standardise between 0 and 1 across parameter space.
Because of the negative relationship found for initial abundance, the standardisation is reversed to make the effect comparable to the other parameters.

```{r}
dataStandard <- data
for(parameter in parameters){
  raw <- data[,parameter]
  standard <- (raw-min(raw))/(max(raw)-min(raw))
  data[,paste0(parameter,".norm")] <- standard
}

data$initialAbundance.norm <- 1-data$initialAbundance.norm

```

Create vectors of pairwise interactions and bind them to the data frame.
Interaction strength is just the mean of the normalised parameter values for each run.

```{r}

combinations <- combn(paste0(parameters,".norm"),2)
interactions.pairwise <- c()
for(i in 1:dim(combinations)[2]){
  
  param1 <- combinations[1,i]
  param2 <- combinations[2,i]
  
  combo <- (data[,param1]+data[,param2])/2
  
  interaction <- paste0(param1,"_X_",param2)
  
  data[,interaction] <- combo
  interactions.pairwise <- c(interactions.pairwise, interaction)
}

```

Create vectors of triple-wise interactions and bind them to the data frame.
Interaction strength is again the mean of the normalised parameter values for each run.

```{r}
combinations <- combn(paste0(parameters,".norm"),3)
interactions.triplewise <- c()
for(i in 1:dim(combinations)[2]){
  
  param1 <- combinations[1,i]
  param2 <- combinations[2,i]
  param3 <- combinations[3,i]
  
  combo <- (data[,param1]+data[,param2]+data[,param3])/3
  
  interaction <- paste0(param1,"_X_",param2,"_X_",param3)
  
  data[,interaction] <- combo
  interactions.triplewise <- c(interactions.triplewise, interaction)
}
```


Explore relationships graphically between pair-/triple-wise interaction values and the richness discrepancy.
Linear regressions are fitted to each relationship.

```{r, warnings = FALSE}

for(interaction in c(interactions.pairwise,interactions.triplewise)){
  plotData <- data.frame(data$richness.discrepancy,data[,interaction])
  
  regression <- lm(plotData[,2] ~ plotData[,1])
  
  print(ggplot(data, aes(x=richness.discrepancy, y=plotData[,2]))+
  labs(y=interaction, x="richness discrepancy") +
    scale_fill_viridis_c() +
    geom_abline(slope = regression$coefficients[2],
                intercept = regression$coefficients[1]) +
  geom_point(aes(fill=species.surviving),colour = "black", shape =21, size = 2) +
  theme_classic())
}


```

Explore relationship of the interaction value of all the parameters and richness discrepancy.

```{r}


all_params <- rowSums(data[,paste0(parameters,".norm")])/4

data$interaction_all_params <- all_params

regression <- lm(data$interaction_all_params ~ data$richness.discrepancy)

  print(ggplot(data, aes(x=richness.discrepancy, y=interaction_all_params))+
  labs(y="interaction_all_params", x="richness discrepancy") +
    scale_fill_viridis_c() +
    geom_abline(slope = regression$coefficients[2],
                intercept = regression$coefficients[1]) +
  geom_point(aes(fill=species.surviving),colour = "black", shape =21, size = 2) +
  theme_classic())


```

























