---
title: "Per cell metric functions"
output: html_document
---

```{r echo = FALSE}
# load functions
path_function <- "C:/Users/thoma/OneDrive/Documents/aynsScreeuyn/genesis/functions"
functions     <- list.files(path = path_function,pattern = ".R")
invisible(lapply(paste(path_function,functions,sep = "/"),source))
```

## All the functions!

```{r echo = FALSE}
lsf.str()
```

## Cluster divergence
Calculate cluster divergence per cell from a presence/absence data frame.

1. For each cell, identify all the species present.

2. For each species, average the divergence between its constituent clusters.

3. For each cell, average the divergences of all species present in that cell (values from 1 and 2).

```{r}
clusterDivergence
```

## Cluster diversity
Calculate cluster divergence per cell from a presence/absence data frame.

1. For each cell, identify all the species present.

2. For each species, calculate the number of clusters they have across their entire range.

3. For each cell, sum the number of constituent clusters for all species present in that cell.

```{r}
clusterDiversity
```

## Calculate trait summary metrics per cell
Calculate summary metrics for a specified trait per cell. This aggregates the traits of all species present in a cell.

These are currently:

* minimum

* maximum

* range

* standard deviation

Traits in the present simulations are:

* thermal optimum

* niche value


```{r}
traitMetricsCell
```

## Trait Variance Cell
Calculates the mean variance of all the mean trait values of each species present in a cell across their entire range.

```{r}
traitVarianceCell
```


## Continuity
Calculates a normalised comparison between two chosen metrics.

1. Remove NA values

2. Normalise the metrics between the the global maximum and global minimum values for that time step.

3. Calculate the square root of the squared difference between the normalised metric values per cell.


```{r}
continuity
```





