---
title: "Exciting Analysis"
output: html_document
---
  
#### Load packages and data.
The data are summary metrics computed by batchSummary.R
Remove runs that resulted in no species.

```{r, message = FALSE}
library(tidyverse)
library(shapper)
library(gridExtra)
library(readxl)

ecoregions           <- read_csv("C:/Users/thoma/OneDrive/Documents/PhD/chapter_1/analysis/regional_continuity/01_sim_ecoregions.csv")
regional_comparisons <- read_csv("C:/Users/thoma/OneDrive/Documents/PhD/chapter_1/analysis/regional_continuity/02_regional_continuity_comparisons.csv")
regional_continuity  <- read_csv("C:/Users/thoma/OneDrive/Documents/PhD/chapter_1/summaries/regional_continuity.csv")
run_summaries        <- read_csv("C:/Users/thoma/OneDrive/Documents/PhD/chapter_1/summaries/all.csv")

# calculate overall run continuity ####

run_summaries$continuity_all <- run_summaries$species.surviving/run_summaries$clusters.total

# make data joinable ####
colnames(regional_continuity)[4]  <- "continuity_regional"
colnames(regional_comparisons)[4] <- "continuity_comparisons"

regional_continuity$id <- paste0(regional_continuity$batch,"_",regional_continuity$run)
regional_continuity    <- regional_continuity[,3:5]

regional_comparisons$id <- paste0(regional_comparisons$batch,"_",regional_comparisons$run)
regional_comparisons    <- regional_comparisons[,3:5]

run_summaries$id       <- paste0(run_summaries$batch_id,"_all_",run_summaries$run_id)

# join the data ####

data_regional   <- inner_join(run_summaries,regional_continuity, by = "id")
data_comparison <- inner_join(run_summaries,regional_comparisons, by = "id")

print("data_regional")
colnames(data_regional)
print("data_comparison")
colnames(data_comparison)

```

#### Plot out continuity values of each region

Calculate the overall continuity of each run and explore the impact of the different parameters.

```{r}

data <- data_regional %>% filter(species.surviving > 5)

parameters <- c("t_opt","dispRange","speciation","initialAbundance")


for(parameter in parameters){
  plotData <- data.frame(data$continuity_all,data[,parameter])
  
  regression <- lm(plotData[,1] ~ plotData[,2])
  
  print(ggplot(data, aes(x=plotData[,2], y=continuity_all))+
    labs(x=parameter, y="continuity_all") +
    ylim(c(0,1)) +
    scale_fill_viridis_c() +
    geom_point(aes(fill=richness.discrepancy),colour = "black", shape =21, size = 2) +
    #geom_abline(slope = regression$coefficients[2],
    #            intercept = regression$coefficients[1]) +
  theme_classic()
  )
}
```

The same again, but split into biological realms as according to Spalding et al. 2007.

```{r}

parameters <- c("t_opt","dispRange","speciation","initialAbundance")


for(parameter in parameters){
  plotData <- data.frame(data$continuity_regional,data[,parameter])
  
  regression <- lm(plotData[,1] ~ plotData[,2])
  
  print(ggplot(data, aes(x=plotData[,2], y=continuity_regional))+
    labs(x=parameter, y="continuity_regional") +
    ylim(c(0,1)) +
    geom_point(aes(fill=realm),colour = "black", shape =21, size = 2) +
  theme_classic()
  )
}

```

#### Regional continuity comparisons against parameter values

```{r}
compare <- unique(data_comparison$comparison)
titles  <- c("Central Indo-Pacific vs Tropical Atlantic",
             "Central Indo-Pacific vs Tropical Eastern Pacific",
             "Central Indo-Pacific vs Western Indo-Pacific",
             "Tropical Atlantic vs Tropical Eastern Pacific",
             "Tropical Atlantic vs Western Indo-Pacific",
             "Tropical Eastern Pacific vs Western Indo-Pacific")

for(i in 1:length(compare)){
  
  data <- data_comparison %>% filter(species.surviving > 5 & comparison == compare[i])
  
  parameters <- c("t_opt","dispRange","speciation","initialAbundance")
  
  
  for(parameter in parameters){
    plotData <- data.frame(data$continuity_comparisons,data[,parameter])
    
    regression <- lm(plotData[,1] ~ plotData[,2])
    
    print(ggplot(data, aes(x=plotData[,2], y=continuity_comparisons))+
      labs(x=parameter, y=compare[i]) +
      ggtitle(titles[i]) +
      ylim(c(0,1)) +
      geom_point() +
      #geom_point(aes(fill=comparison),colour = "black", shape =21, size = 2) +
      #geom_abline(slope = regression$coefficients[2],
      #            intercept = regression$coefficients[1]) +
    theme_classic()
    )
  }
}


```



