---
title: "Parameter Exploration"
output: html_document
---
<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 10000)
```

There are 2 measures of continuity to be explored:  
  
* **contStructure**  is total extant species / total extant clusters.  
* **contDivergence** is MPD between extant species / MPD between extant clusters.


There are 5 parameters:

* **t_opt**:        the initial thermal optimum of the first species in a simulation.  
* **dispRange**:    the dispersal range of all species in a simulation.  
* **speciation**:   the distance between clusters at which a new species is formed.  
* **mutationRate**: the standard deviation of the gaussian distribution which is used to mutate all traits.  
* **n_width**:      the competitive niche width of all species in the simulation. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(MuMIn)
library(gridExtra)

# load functions
path_function <- "C:/Users/thoma/OneDrive/Documents/aynsScreeuyn/genesis/functions"
functions     <- list.files(path = path_function,pattern = ".R")
invisible(lapply(paste(path_function,functions,sep = "/"),source))

# load data
data <- read_csv("C:/Users/thoma/OneDrive/Documents/PhD/chapter_1/summaries/6.2_all.csv")
data <- data %>% filter(species.surviving > 9)
data <- data %>% filter(species.total < 1000)

# calculate metrics
data$contStructure <- data$species.surviving/data$clusters.total
data$contDivergence <- data$species.divergence/data$clusters.divergence.sim
data$strVdiv   <- continuity(data$contStructure,data$contDivergence)

data$species.turnover <- data$species.surviving/data$species.total
```

# 1. Results
## 1.1 Parameters vs contintuity
### 1.1.1 Visualise individual relationships

```{r, echo = FALSE, message = FALSE, tidy = TRUE}
parameters <- c("t_opt",
                "dispRange",
                "speciation",
                "mutationRate",
                "n_width")

conts <- c("contStructure",
           "contDivergence")

plotData_1 <- pivot_longer(data, cols = c(contDivergence,
                                        contStructure))
grobs <- list()
plot_no <- 1
for(c in 1:length(conts)){
  
  plotData <- plotData_1 %>% filter(name == conts[c])

  for(i in 1:length(parameters)){
    
    plotData[,"parameter"] <- plotData[,parameters[i]]
    
    plot <- ggplot(plotData, aes(x=parameter,
                                 y=value)) +
      geom_point() +
      ylab(conts[c]) +
      xlab(parameters[i]) +
      stat_smooth()
    
    grobs[[plot_no]] <- plot
  plot_no <- plot_no +1
  }
}

do.call("grid.arrange", c(grobs, ncol=5))
```


### 1.1.2 Simple additive multiple linear regression

```{r message=FALSE, warning=FALSE, echo = FALSE}

modelData <- data
polynominal <- 2

models <- list()
for(c in 1:length(conts)){
  
  modelData[,"response"] <- modelData[,conts[c]]
  
  model <- lm(response ~ 
                poly(t_opt,polynominal) + 
                poly(dispRange,polynominal) + 
                poly(speciation,polynominal) + 
                poly(mutationRate,polynominal) + 
                poly(n_width,polynominal),
              data = modelData,
              na.action = "na.fail")
  
  models[[conts[c]]] <- model
}

structure_dredged  <- dredge(models[[1]])
divergence_dredged <- dredge(models[[2]])

```

**Structure continuity:**
```{r message=FALSE, warning=FALSE}
summary(models$contStructure)
```

**Divergence continuity:**
```{r message=FALSE, warning=FALSE}
summary(models$contDivergence)
```
Retaining third level polynomials doesn't seem realistic, so will retain up to second level for dispRange and speciation.  

### 1.1.3 Multiple linear regression with filtered terms and interactions
```{r message=FALSE, warning=FALSE, echo = FALSE}

modelData <- data
polynominal <- 2

models <- list()
for(c in 1:length(conts)){
  
  modelData[,"response"] <- modelData[,conts[c]]
  
  model <- lm(response ~ 
                poly(dispRange,polynominal) * 
                poly(speciation,polynominal),
              data = modelData,
              na.action = "na.fail")
  
  models[[conts[c]]] <- model
}

```
**Structure continuity:**
```{r message=FALSE, warning=FALSE}
summary(models$contStructure)
```
**Divergence continuity:**
```{r message=FALSE, warning=FALSE}
summary(models$contDivergence)
```

Next steps:
* plot model prediction values over data  
* generate model predictions holding parameters constant (min, max, mean)



## Parameters vs components of continuity









## Explore continuity vs metrics

```{r}
metrics <- c("species.total",
             "species.surviving",
             "species.turnover",
             "species.divergence",
             "species.range",
             "t_opt.evenness",
             "t_opt.diversity",
             "t_opt.mean",
             "niche.mean",
             "niche.evenness",
             "niche.diversity",
             "niche.sd",
             "richness.avg",
             "richness.sd",
             "richness.discrepancy",
             "clusters.total",
             "clusters.divergence.sim")


plotData <- data

for(c in 1:length(conts)){
  
  plotData[,"contintuity"] <- plotData[,conts[c]]
  
  for(i in 1:length(metrics)){
    
    
    plotData[,"metric"] <- plotData[,metrics[i]]
    
    plot <- ggplot(plotData) +
      geom_point(aes(x=metric,
                     y=continuity)) +
      ylab(conts[c]) +
      xlab(metrics[i])
    
    print(plot)
    
  }
}
```

## Explore parameters vs metrics


```{r}


plotData <- data

for(p in 1:length(parameters)){
  
  plotData[,"parameter"] <- plotData[,parameters[p]]
  
  for(i in 1:length(metrics)){
    
    plotData[,"metric"] <- plotData[,metrics[i]]
    
    plot <- ggplot(plotData) +
      geom_point(aes(x=parameter,
                     y=metric)) +
      ylab(metrics[i]) +
      xlab(parameters[p])
    
    print(plot)
    
  }
  
}
```








































